Meta-analytic findings of the self-controlled motor learning literature: Underpowered, biased, and lacking evidential value

### 2021-08-25 docs/revision/manuscript-revised.Rmd
update: revisions based on round 1 comments

keywords          : "Motor learning, retention, choice, ``OPTIMAL'' theory, meta-analysis, p-curve, publication bias"
# wordcount         : "5055"
bibliography      : ["../refs.bib"]
annotate_references: yes
nocite: |
  @Aiken2012-po,@Alami2013-gy,@Ali2012-rq,@Andrieux2016-lf,@Andrieux2012-dj,@Arsal2004-kf,@Barros2010-pc,@Barros2019-my,@Bass2015-cc,@Bass2018-vz,@Brydges2009-zb,@Bund2004-hy,@Carter2012-sj,@Chen2002-vg,@Chiviacowsky2014-ob,@Chiviacowsky2017-fy,@Chiviacowsky2002-ep,@Chiviacowsky2012-pk,@Chiviacowsky2008-bj,@Chiviacowsky2012-ri,@davis2009-dt,@Fagundes2013-ge,@Fairbrother2012-gz,@Ferreira2019-zo,@Figueiredo2018-kw,@Ghorbani2019-yz,@Grand2015-de,@Grand2017-de,@Hansen2011-rr,@Hartman2007-uv,@Hemayattalab2013-mq,@Ho2016-on,@Holmberg2013-qk,@Huet2009-wu,@Ikudome2019-ru,@Jalalvand2019-im,@Janelle1997-ht,@jones2010-qw,@Kaefer2014-bs,@Keetch2007-yp,@Kim2019-sl,@Leiker2016-zd,@Leiker2019-fz,@Lemos2017-qx,@Lessa2015-eq,@Lewthwaite2015-bd,@Lim2015-qs,@Marques2016-aw,@Marques2017-ue,@Norouzi2016-cx,@Nunes2019-nr,@Ostrowski2015-nb,@Patterson2010-tk,@Patterson2010-uu,@Patterson2013-nh,@Patterson2011-vt,@Post2016-vg,@Post2011-qc,@Post2014-yi,@Rydberg2011-zg,@Sanli2013-qh,@Ste-Marie2013-uc,@Tsai2015-rj,@Von_Lindern2017-al,@Williams2017-pj,@Wu2011-ha,@wu2007self,@Wulf2014-sn,@Wulf1999-pn,@Wulf2001-nb,@Wulf2018-ie,@Wulf2005-sz

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(papaja)
library(kableExtra)
library(tidyverse)
# r_refs("../r-references.bib")
# Read in data files for tables
table_1 <- readr::read_csv("../revision/table-1.csv", show_col_types = FALSE)
table_a1 <- readr::read_csv("../revision/table-a1.csv", show_col_types = FALSE)
```

@@

A contrasting line of research has been reported by Carter and his colleagues [-@Carter2014-up; -@Carter2017-ix; -@Carter2017-mk] in which informational factors, the second dominant perspective, are given more weight as an explanatory variable. In one experiment @Carter2014-up, self-control participants were provided with choice over receiving knowledge of results, but divided into three experimental groups; those who could make their knowledge of results decision: before the trial, after the trial, or both (they would decide before, but could change their mind following the trial). Timing of the choice significantly attenuated the self-control benefit. While the self-after and self-both groups exhibited learning advantages relative to their yoked counterparts, the self-before group displayed no such advantage. The argument proffered by the researchers was that there was more informational value to be gained from knowledge of results requested after a trial than when it had to be requested before the outcome of the trial occurred [also see @Chiviacowsky2005-iq].

@@

In the present research, these different viewpoints concerning the mechanisms of self-controlled learning advantages were examined via meta-analysis with choice-type included as a moderator. The logic was that the motivational and informational perspectives would have different predictions. More specifically, from a motivation hypothesis, no moderating effect of choice-type on motor learning would be expected. In contrast, smaller effects for irrelevant-choice type, as compared to relevant-choice types would be expected from the information-processing perspective.

@@

In sum, the objective of this meta-analysis was to estimate the true average effect of self-controlled learning and evaluate the evidential value of the self-controlled learning literature. Bias resulting from selective publication was addressed with weight function and *p*-curve models and effect size estimates were adjusted accordingly. A key theoretical question was also addressed through moderator analyses, but, to anticipate, inferences will depend on the reliability of the evidence overall. Finally, sensitivity analyses were conducted in addition to pre-registered analyses in an effort to understand the extent that our conclusions depended on the modelling techniques and assumptions adopted.

# Methods

## Pre-registration
The procedures followed to conduct this meta-analysis were pre-registered and can be viewed at https://osf.io/qbg69 \hyperref[sec:sharing]{(see Data, materials, and code availability section)}. This meta-analysis was retrospective and earlier samples of the literature had been meta-analyzed prior to this pre-registration, albeit with different data collection procedures, scope, and excluding recent experiments.

@@

The search commenced at PubMed and Google Scholar with the following query: *self-control** OR *self-regulat** OR *self-direct** OR *learner-control** OR *learner-regulat** OR *learner-direct** OR *subject-control** OR *subject-regulat** OR *subject-direct** OR *performer-control** OR *performer-regulat** OR *performer-direct** AND *motor learning**. The query retrieved 9014 hits on PubMed and 98,600 hits on Google Scholar. Each researcher excluded hits based on title alone or title and abstract when necessary, and quit searching the databases at self-selected intervals following extended periods of excluding 100\% of search results. Following an initial run of searching databases, each researcher employed their own search strategies, including reviewing the reference sections of reviews and included articles, consulting the ``OPTIMAL'' theory website[^1], and searching the ProQuest Thesis database.

@@

This literature search process resulted in 160 articles that could not be excluded without consulting the full-text of the article. All 160 articles were coded for inclusion or exclusion by two researchers independently. All instances of disagreement between coders were reviewed by three authors (BM, ZY, and JH), and consensus was reached in each case. Disagreements were infrequent and were often caused by a lack of clarity in the articles (e.g., 100% knowledge of results groups labelled as yoked groups). None of the coding disagreements evolved into conceptual disagreements. Rather, in each case, it was identified that one coder had missed a detail in the full text that changed its inclusion eligibility. Subsequent to this process, a total of 73 articles, which included 78 experiments, met the inclusion criteria (see Table \@ref(tab:table1)).

## Dependent variable selection
The focus of this meta-analysis was on performance outcomes associated with the goal of the skill. The primary theoretical perspectives offered as an account for self-controlled learning are likewise focused on performance outcomes. For example, the ``OPTIMAL'' theory proposes that a learner's movements become coupled with the goal they are trying to achieve when they experience autonomy-support during practice [@Wulf2016-gf]. To reflect this focus, a dependent measure priority list was developed that gave higher priority to absolute error measures and less priority to consistency measures, time/work measures, and form scores. Dependent measure priority was ordered as follows: 1) absolute error (and analogous measures: radial error, points in an accuracy measure), 2) root-mean-square-error (RMSE), 3) absolute constant error, 4) variable error, 5), movement time (and distance travelled), 6) movement form – expert raters, 7) otherwise unspecified objective performance measure reported first in research report.[^2] In the event that multiple measures of motor performance were reported for an experiment, effect sizes were calculated for the highest priority measure reported in the study. In experiments with multiple self-control groups and one yoked group, the self-control groups were combined [@Higgins2011-rg]. If multiple choice-types or sub-populations were included in an experiment, combined and individual effects were calculated for inclusion in moderator analyses.

@@

@@

Bias correction methods vary in their performance depending on the total amount of heterogeneity, the true average effect size, the amount of publication bias, and the intensity of *p*-hacking in the data [@Carter2019-vv]. To determine which bias correction models perform well in the various plausible conditions for data in this meta-analysis, model performance checks were conducted using the \href{http://www.shinyapps.org/apps/metaExplorer/}{Meta-Showdown Explorer} shiny app developed by Carter and colleagues [-@Carter2019-vv]. Simulated conditions were as follows: Medium publications bias (significant results published at 100\% frequency, non-significant published at 20\% frequency, wrong direction effects published at 5\% frequency), \textcolor{red}{medium questionable research practice environment (QRP; see Carter et al.} [-@Carter2019-vv] \textcolor{red}{for detailed explanation of QRP environment)}, $\tau$ = 0, .2; $g$ = 0, .2, .5; $k$ = 60, good performance defined as a maximum of .1 upward or downward bias, and maximum mean absolute error of .1, also tested with maximum bias and error values of .15. With good performance defined by a maximum bias in either direction of .1 and maximum absolute error of .1, the weight function model and, \textcolor{red}{to a lesser extent,} *p*-curve models provided coverage across all plausible conditions except the highest heterogeneity condition ($\tau$ = .4). With good performance defined as a maximum bias and error of .15, the precision-effect with standard error (PEESE) method provided good performance in all conditions. Therefore, sensitivity analyses were conducted on effect size data via *p*-curve and PEESE methods. \textcolor{red}{An additional sensitivity analysis of the estimated power among included studies was conducted with the} *z*\textcolor{red}{-curve} [@bartovs2020]. *Z*\textcolor{red}{-curve, like} *p*\textcolor{red}{-curve, analyzes only statistically signficant results and estimates the power of the included studies (called expected replication rate, ERR). However, unlike} *p*\textcolor{red}{-curve,} *z*\textcolor{red}{-curve is robust to heterogeneity because it fits a finite mixture model of seven distributions, allowing the underlying true effects to vary. Further,} *z*\textcolor{red}{-curve also estimates the power of all studies that have been conducted (called expected discovery rate, EDR) which can be compared to the observed discovery rate in order to test for the presence of publication bias.}

@@

## Outlier removal
Two studies were flagged as significantly influential outliers by all nine influence metrics calculated during data screening: @Lemos2017-qx, $g$ = 3.7, and @Marques2017-ue, $g$ = 3.95. No other effect sizes were identified as outliers by any metric. Both outliers were removed from all subsequent analyses.

(ref:fig2-caption) Forest plot of Hedges' $g$ (95% CI) for self-controlled versus yoked groups on retention tests. Size of squares is proportional to $1/\sigma^2$ (precision). Light grey polygons represent \textcolor{red}{95\% $CI$ estimates from publication-status moderator analysis. Estimates from unpublished studies center on $g$ = .003 and published studies on $g$ = .54}

@@

## Naïve random effects model
The naïve random effects model estimated the average treatment effect of self-controlled practice, $g$ = .44 ($k$ = 52, $N$ = 3134, 95% $CI$ [.31, .56]).  However, there was significant variability in the average effect estimated across experiments, $Q$($df$ = 51) = 103.45, $p$ < .0001, $\tau$ = .31. It was estimated that 47.9% ($I^2$) of the total variability in effect sizes across experiments was due to true heterogeneity in the underlying effects measured (see Figure \@ref(fig:fig2)).

## Moderator analyses
Six moderators selected for theoretical and/or methodological reasons were tested separately. Five moderators failed to account for a significant amount of heterogeneity: experimental setting ($p$ = .46, $R^2$ = 1%), compensation ($p$ = .99, $R^2$ = 0%), choice-type ($p$ = .71, $R^2$ = 0%), sub-population ($p$ = .74, $R^2$ = 0%), and retention interval ($p$ = .54, $R^2$ = 0%). One moderator, publication status, accounted for a statistically significant amount of heterogeneity, $p$ < .0001, $R^2$ = 48%. Among published experiments self-controlled practice had a strong benefit, $g$ = .54, 95% $CI$ [.28, .81]. However, among unpublished experiments self-controlled practice had essentially no effect, $g$ = .003, 95% $CI$ [-.23, 24].

## Selection model
The weight-function model combines an effect size model and a selection model [@Hedges1996-yh]. The effect size model is equivalent to the naïve random effects model, specifying what the distribution of effect sizes would be in the absence of publication bias or other selection effects. The selection model accounts for the probability a given study survives selection based on its *p*-value and specifies how the effect size distribution is modified by selection. A weight-function model with a *p*-value cutpoint of (one-tailed) .025 was fit to the retention effect size estimates. The results of a likelihood ratio test suggest the adjusted model was a significantly better fit to the data than the unadjusted model, $\chi^2$($df$ = 1) = 21.18, $p$ < .0001.[^4] The adjusted effect size estimate was significantly different from zero, $g$ = .107, $p$ < .001, 95% $CI$ [.05, .17]. According to the adjusted model, non-significant results were 6% as likely to survive selection as significant results (see Figure \@ref(fig:fig3)). \textcolor{red}{Note that the \texttt{weightr} function failed to estimate the random effects model and the results reported here are based on a fixed-effect estimate.}

@@

(ref:fig3-caption) Funnel plot of self-controlled learning studies at retention. Standard error is plotted on the y-axis and Hedges' g is plotted on the x-axis.  Red line represents naïve random effects model estimate (g = .44); blue line represents weight-function model adjusted estimate (g = .11); black contour lines represent two-tailed p-values of .05. In the absence of bias (and other forms of heterogeneity), the most precise experiments would centre on the naïve random effects estimate near the bottom of the plot and as experiments get progressively less precise they would move up the plot and spread out symmetrically. In the presence of bias, one would expect experiments to track to the right side of the positive contour line. The clustering of experiments just to the right of the positive contour line in the above plot suggests substantial bias.

@@

```{r fig3, echo = FALSE, fig.cap = "(ref:fig3-caption)", fig.align = "center"}
knitr::include_graphics("../../figs/fig3.pdf")
```

## *P*-curve
The purpose of the *p*-curve analysis was to investigate the evidential value in the published reports ($N$ = 26) of statistically significant self-controlled learning benefits. Visual inspection of Figure \@ref(fig:fig4) reveals a v-shaped distribution with the greatest frequency of *p*-values in the <.05 bin. The observed *p*-curve was significantly flatter than would be expected if the experiments had 33% power, $p$ = .0035, indicating an absence of evidential value. Conversely, the half *p*-curve [@Simonsohn2015-mn] was significantly right skewed, suggesting the presence of evidential value. Sensitivity analysis, however, revealed that the half curve does not remain significantly right skewed following removal of the most extreme *p*-value from the sample. \textcolor{red}{The estimated power of the included studies was 5\%, 95\% CI [5\%, 17\%].}

## Interim discussion
The primary results described above suggest that selection effects have caused a seriously distorted record of self-controlled learning. Estimated benefits are less than one third of the naïve estimate, $g$ = .107, 95% $CI$ \textcolor{red}{[.05, .17]}. The *p*-curve analysis failed to detect robust evidence of a self-controlled learning effect. The performance of the weight-function model depends on the specific conditions present in the meta-analysis, although these conditions are unknowable [@Carter2019-vv]. It was necessary to conduct sensitivity analyses with additional bias correction methods to assess the reliability of the selection-adjusted weight-function model estimate. Based on performance checks conducted under a range of plausible conditions, it was determined that sensitivity analyses conducted with a PEESE meta-regression and *p*-curve effect size estimation would provide good performance coverage across most plausible conditions.

(ref:fig4-caption) *P*-curve analysis of published experiments that were statistically significant at retention. If the included experiments are studying a true null hypothesis the expected distribution of p-values is uniform, represented by the dotted line. If the experiments are studying a true effect, the expected distribution becomes increasingly right skewed as a function of statistical power. The expected right skewed distribution associated with 33% power is plotted by the dashed line. The observed p-curve is plotted by the solid line and was substantially flatter than the 33% power distribution. The half p-curve analysis included p-values below p = .025 and was significantly right skewed. The right skew did not survive deletion of the most extreme value.

```{r fig4, echo = FALSE, fig.cap = "(ref:fig4-caption)", fig.align = "center", out.width="90%"}
knitr::include_graphics("../../figs/fig4.pdf")
```

@@

## Precision-effect with standard error (PEESE) model
When publication bias is present in a body of evidence, sample size and effect size can be negatively correlated [@Stanley2014-xl]. The PEESE model fits a quadratic relationship between effect size and standard error to reflect the intuition that publication bias is stronger for low precision studies than high precision studies. The rationale is that low precision studies need to overestimate effects to achieve significance and get published, while high precision studies can publish without exaggerated effects; thus, creating greater publication bias among lower precision studies [@Carter2019-vv; @Stanley2014-xl]. A weighted-least-squares regression model was fit with effect size regressed on the square of the standard error, weighted by the inverse of the variance:
$$g_{i} = b_{0} + b_{1}se_{i}^2 + e_{i}$$
The PEESE method estimated a non-significant benefit of self-controlled learning after controlling for publication bias, $g$ = .054, $p$ = .659.

## *P*-curve effect estimation
A *p*-curve model was fit to the overall retention effect size data, unlike the first primary *p*-curve which was fit to the reported significant results. The *p*-curve is a function of sample size and effect size, and because sample size is known, the effect size that provides the best fit to the observed *p*-curve can be estimated [@Simonsohn2014-xq]. A *p*-curve analysis conducted with th`R` package `dmetar` [@Harrer2019-yv] was used to estimate the average effect size among the statistically significant effects in the meta-analysis. The model estimated an average effect of $g$ = .035.[^5] \textcolor{red}{The estimated power of included studies was 7\%, 95\% $CI$ [5\%, 22\%]. Unfortunately,} *p*\textcolor{red}{-curve does not perform well in the presence of heterogeneity and these results should be interpreted cautiously.}

[^5]: The *p*-curve of effect sizes was significantly flatter than the expected 33% power curve as well, $p$ = .009.

>>>>

keywords          : "Motor learning, retention, choice, ``OPTIMAL'' theory, meta-analysis, p-curve, publication bias"
# wordcount         : "X"
bibliography      : ["../refs.bib"]
annotate_references: yes
nocite: |
  @Aiken2012-po,@Alami2013-gy,@Ali2012-rq,@Andrieux2016-lf,@Andrieux2012-dj,@Arsal2004-kf,@Barros2010-pc,@Barros2019-my,@Bass2015-cc,@Bass2018-vz,@Brydges2009-zb,@Bund2004-hy,@Carter2012-sj,@Chen2002-vg,@Chiviacowsky2014-ob,@Chiviacowsky2017-fy,@Chiviacowsky2002-ep,@Chiviacowsky2012-pk,@Chiviacowsky2008-bj,@Chiviacowsky2012-ri,@davis2009-dt,@Fagundes2013-ge,@Fairbrother2012-gz,@Ferreira2019-zo,@Figueiredo2018-kw,@Ghorbani2019-yz,@Grand2015-de,@Grand2017-de,@Hansen2011-rr,@Hartman2007-uv,@Hemayattalab2013-mq,@Ho2016-on,@Holmberg2013-qk,@Huet2009-wu,@Ikudome2019-ru,@Jalalvand2019-im,@Janelle1997-ht,@jones2010-qw,@Kaefer2014-bs,@Keetch2007-yp,@Kim2019-sl,@Leiker2016-zd,@Leiker2019-fz,@Lemos2017-qx,@Lessa2015-eq,@Lewthwaite2015-bd,@Lim2015-qs,@Marques2016-aw,@Marques2017-ue,@Norouzi2016-cx,@Nunes2019-nr,@Ostrowski2015-nb,@Patterson2010-tk,@Patterson2010-uu,@Patterson2013-nh,@Patterson2011-vt,@Post2016-vg,@Post2011-qc,@Post2014-yi,@Rydberg2011-zg,@Sanli2013-qh,@Ste-Marie2013-uc,@Tsai2015-rj,@Von_Lindern2017-al,@Williams2017-pj,@Wu2011-ha,@wu2007self,@Wulf2014-sn,@Wulf1999-pn,@Wulf2001-nb,@Wulf2018-ie,@Wulf2005-sz,@wulf2015-mg

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(papaja)
library(kableExtra)
library(tidyverse)
# r_refs("../r-references.bib")
# Read in data files for tables
table_1 <- readr::read_csv("../revision/table-1.csv")
table_a1 <- readr::read_csv("../revision/table-a1.csv")
```

@@

A contrasting line of research has been reported by Carter and his colleagues [-@Carter2014-up; -@Carter2017-ix; -@Carter2017-mk] in which informational factors, the second dominant perspective, are given more weight as an explanatory variable. In one experiment by @Carter2014-up, self-control participants were provided with choice over receiving knowledge of results, but divided into three experimental groups; those who could make their knowledge of results decision: before the trial, after the trial, or both (they would decide before, but could change their mind following the trial). Timing of the choice significantly attenuated the self-control benefit. While the self-after and self-both groups exhibited learning advantages relative to their yoked counterparts, the self-before group displayed no such advantage. The argument proffered by the researchers was that there was more informational value to be gained from knowledge of results requested after a trial than when it had to be requested before the outcome of the trial occurred [also see @Chiviacowsky2005-iq].

@@

In the present research, these different viewpoints concerning the mechanisms of self-controlled learning advantages were examined via meta-analysis with choice-type included as a moderator. The logic was that the motivational and informational perspectives would have different predictions. More specifically, from a motivation hypothesis, no moderating effect of choice-type on motor learning would be expected. In contrast, smaller effects for irrelevant-choice type, as compared to relevant-choice types, would be expected from the information-processing perspective.

@@

In sum, the objectives of this meta-analysis were to estimate the true average effect of self-controlled learning and evaluate the evidential value of the self-controlled learning literature. Bias resulting from selective publication was addressed with weight function and *p*-curve models and effect size estimates were adjusted accordingly. A key theoretical question \textcolor{red}{related to the underlying mechanisms of putative self-controlled learning advantages (motivational versus informational influences)} was also addressed through moderator analyses, but, to anticipate, inferences will depend on the reliability of the evidence overall. Finally, sensitivity analyses were conducted in addition to pre-registered analyses in an effort to understand the extent that our conclusions depended on the modeling techniques and assumptions adopted.

# Method

## Pre-registration
The procedures followed to conduct this meta-analysis were pre-registered and can be viewed at https://osf.io/qbg69 \hyperref[sec:sharing]{(see Data, materials, and code availability section)}. This meta-analysis was retrospective and earlier samples of the literature had been meta-analyzed prior to this pre-registration, albeit with different data collection procedures, scope, and excluding recent experiments. \textcolor{red}{This study adheres to PRISMA reporting guidelines} [@page2021].

@@

The search commenced at PubMed and Google Scholar with the following query: *self-control** OR *self-regulat** OR *self-direct** OR *learner-control** OR *learner-regulat** OR *learner-direct** OR *subject-control** OR *subject-regulat** OR *subject-direct** OR *performer-control** OR *performer-regulat** OR *performer-direct** AND *motor learning**. The query retrieved 9014 hits on PubMed and 98,600 hits on Google Scholar. Each researcher excluded hits based on title alone or title and abstract when necessary, and quit searching the databases at self-selected intervals following extended periods of excluding 100\% of search results. Following an initial run of searching databases, each researcher employed their own search strategies, including reviewing the reference sections of reviews and included articles, consulting the "OPTIMAL" theory website[^1], and searching the ProQuest Thesis database.

@@

This literature search process resulted in 160 articles that could not be excluded without consulting the full-text of the article. All 160 articles were coded for inclusion or exclusion by two researchers independently. All instances of disagreement between coders were reviewed by three authors (BM, ZY, and JH), and consensus was reached in each case. Disagreements were infrequent and were often caused by a lack of clarity in the articles (e.g., 100% knowledge of results groups labeled as yoked groups). None of the coding disagreements evolved into conceptual disagreements. Rather, in each case, it was identified that one coder had missed a detail in the full text that changed its inclusion eligibility. Subsequent to this process, a total of 73 articles, which included 78 experiments, met the inclusion criteria (see Table \@ref(tab:table1)).

## Dependent variable selection
The focus of this meta-analysis was on performance outcomes associated with the goal of the skill. The primary theoretical perspectives offered as an account for self-controlled learning are likewise focused on performance outcomes. For example, the "OPTIMAL" theory proposes that a learner's movements become coupled with the goal they are trying to achieve when they experience autonomy-support during practice [@Wulf2016-gf]. To reflect this focus, a dependent measure priority list was developed that gave higher priority to absolute error measures and less priority to consistency measures, time/work measures, and form scores. Dependent measure priority was ordered as follows: 1) absolute error (and analogous measures: radial error, points in an accuracy measure), 2) root-mean-square-error (RMSE), 3) absolute constant error, 4) variable error, 5), movement time (and distance travelled), 6) movement form – expert raters, 7) otherwise unspecified objective performance measure reported first in research report.[^2] In the event that multiple measures of motor performance were reported for an experiment, effect sizes were calculated for the highest priority measure reported in the study. In experiments with multiple self-control groups and one yoked group, the self-control groups were combined [@Higgins2011-rg]. If multiple choice-types or sub-populations were included in an experiment, combined and individual effects were calculated for inclusion in moderator analyses.

@@

## \textcolor{red}{Risk of bias}
\textcolor{red}{All articles were assessed for risk of bias by the lead author using the Cochrane Risk of Bias 1.0 tool} [@higgins2011]. \textcolor{red}{Each article was coded as either high risk, unclear (some concerns), or low risk on 7 dimensions: sequence generation, allocation concealment, incomplete outcome data, selective outcome reporting, blinding of outcome assessment, blinding of participants and personnel, and other sources of bias.}

@@

Bias correction methods vary in their performance depending on the total amount of heterogeneity, the true average effect size, the amount of publication bias, and the intensity of *p*-hacking in the data [@Carter2019-vv]. To determine which bias correction models perform well in the various plausible conditions for data in this meta-analysis, model performance checks were conducted using the \href{http://www.shinyapps.org/apps/metaExplorer/}{Meta-Showdown Explorer} shiny app developed by Carter and colleagues [-@Carter2019-vv]. Simulated conditions were as follows: Medium publications bias (significant results published at 100\% frequency, non-significant published at 20\% frequency, wrong direction effects published at 5\% frequency), \textcolor{red}{medium questionable research practice environment (QRP; see Carter et al.} [-@Carter2019-vv] \textcolor{red}{for detailed explanation of QRP environment)}, $\tau$ = 0, .2; $g$ = 0, .2, .5; $k$ = 60, good performance defined as a maximum of .1 upward or downward bias, and maximum mean absolute error of .1, also tested with maximum bias and error values of .15. With good performance defined by a maximum bias in either direction of .1 and maximum absolute error of .1, the weight function model and, \textcolor{red}{to a lesser extent,} *p*-curve models provided coverage across all plausible conditions except the highest heterogeneity condition ($\tau$ = .4). With good performance defined as a maximum bias and error of .15, the precision-effect with standard error (PEESE) method provided good performance in all conditions. Therefore, sensitivity analyses were conducted on effect size data via *p*-curve and PEESE methods. \textcolor{red}{An additional sensitivity analysis of the estimated power among included studies was conducted with the \emph{z}-curve [@bartovs2020]. \emph{Z}-curve, like \emph{p}-curve, analyzes only statistically significant results and estimates the power of the included studies (called expected replication rate, ERR). However, unlike \emph{p}-curve, \emph{z}-curve is robust to heterogeneity because it fits a finite mixture model of seven distributions, allowing the underlying true effects to vary. Further, \emph{z}-curve also estimates the power of all studies that have been conducted (called expected discovery rate, EDR) which can be compared to the observed discovery rate in order to test for the presence of publication bias.}

@@

## \textcolor{red}{Risk of bias}
\textcolor{red}{The risk of bias assessment revealed lackluster reporting standards were pervasive among the included articles} (see Figure \@ref(fig:fig2)). \textcolor{red}{For example, comparing a self-control group to a yoked group usually involves first collecting a self-control participant, then their yoked counterpart. Despite this, most articles simply reported that the participants were randomly assigned to these conditions, with no indication of how this temporal constraint was addressed. A similar issue was observed with respect to addressing outliers and attrition. Over 75\% of the included articles failed to mention outliers and how they were addressed (captured by the incomplete outcome data dimension). Most studies included in this study were not double-blind, largely due to the inherent difficulties in conducting a double-blind study of self-controlled motor learning. While the risk of bias associated with a lack of double blinding has been debated} [see @howick2008], \textcolor{red}{it is nonetheless notable that double-blinding was rare among the included studies.}

(ref:fig2-caption) \textcolor{red}{Proportion of studies with low risk, some concerns, and high risk of bias in each of the seven dimensions of the Cochrane RoB 1.0 tool.}

@@

## Outlier removal
Two studies were flagged as significantly influential outliers by all nine influence metrics calculated during data screening: @Lemos2017-qx, *g* = 3.7, and @Marques2017-ue, *g* = 3.95. No other effect sizes were identified as outliers by any metric. Both outliers were removed from all subsequent analyses.

(ref:fig3-caption) Forest plot of Hedges' *g* (95% CI) for self-controlled versus yoked groups on retention tests. Size of squares is proportional to $1/\sigma^2$ (precision). Light grey polygons represent \textcolor{red}{95\% $CI$ estimates from publication-status moderator analysis. Estimates from unpublished studies center on $g$ = .003 and published studies on $g$ = .54}

```{r fig3, echo = FALSE, fig.cap = "(ref:fig3-caption)", fig.align = "center", out.height = "81%"}
knitr::include_graphics("../../figs/fig3.pdf")
```

## Naïve random effects model
The naïve random effects model estimated the average treatment effect of self-controlled practice, *g* = .44 (*k* = 52, *N* = 3134, 95% *CI* [.31, .56]).  However, there was significant variability in the average effect estimated across experiments, *Q*(*df* = 51) = 103.45, *p* < .0001, $\tau$ = .31. It was estimated that 47.9% ($I^2$) of the total variability in effect sizes across experiments was due to true heterogeneity in the underlying effects measured (see Figure \@ref(fig:fig3)).

## Moderator analyses
Six moderators selected for theoretical and/or methodological reasons were tested separately. Five moderators failed to account for a significant amount of heterogeneity: experimental setting (*p* = .46, $R^2$ = 1%), compensation (*p* = .99, $R^2$ = 0%), choice-type (*p* = .71, $R^2$ = 0%), sub-population (*p* = .74, $R^2$ = 0%), and retention interval (*p* = .54, $R^2$ = 0%). One moderator, publication status, accounted for a statistically significant amount of heterogeneity, *p* < .0001, $R^2$ = 48%. Among published experiments, self-controlled practice had a strong benefit, *g* = .54, 95% *CI* [.28, .81]. However, among unpublished experiments, self-controlled practice had essentially no effect, *g* = .003, 95% *CI* [-.23, 24].

## Selection model
The weight-function model combines an effect size model and a selection model [@Hedges1996-yh]. The effect size model is equivalent to the naïve random effects model, specifying what the distribution of effect sizes would be in the absence of publication bias or other selection effects. The selection model accounts for the probability a given study survives selection based on its *p*-value and specifies how the effect size distribution is modified by selection. A weight-function model with a *p*-value cutpoint of (one-tailed) .025 was fit to the retention effect size estimates(see Figure \@ref(fig:fig4)). The results of a likelihood ratio test suggest the adjusted model was a significantly better fit to the data than the unadjusted model, $\chi^2$(*df* = 1) = 21.18, *p* < .0001.[^4] The adjusted effect size estimate was significantly different from zero, *g* = .107, *p* < .001, 95% *CI* [.05, .17]. According to the adjusted model, non-significant results were 6% as likely to survive selection as significant results. \textcolor{red}{Note that the \texttt{weightr} function failed to estimate the random effects model and the results reported here are based on a fixed-effect estimate.}

@@

(ref:fig4-caption) Funnel plot of self-controlled learning studies at retention. Standard error is plotted on the y-axis and Hedges' g is plotted on the x-axis. \textcolor{red}{Dark gray contour regions represent two-tailed p-values between .10 and .05 (not quite significant). The light gray contour regions represent two-tailed p-values between .05 and .01.} In the absence of bias (and other forms of heterogeneity), the most precise experiments would centre on the naïve random effects estimate near the \textcolor{red} {top} of the plot and as experiments get progressively less precise they would move \textcolor{red}{down} the plot and spread out symmetrically. In the presence of bias, one would expect experiments to \textcolor{red} cluster in the light gray contour regions. The clustering of experiments \textcolor{red}{in the positive light gray  contour region} in the above plot suggests substantial bias.

@@

```{r fig4, echo = FALSE, fig.cap = "(ref:fig4-caption)", fig.align = "center"}
knitr::include_graphics("../../figs/fig4.pdf")
```

## *P*-curve
The purpose of the *p*-curve analysis was to investigate the evidential value in the published reports (*N* = 26) of statistically significant self-controlled learning benefits. Visual inspection of Figure \@ref(fig:fig4) reveals a v-shaped distribution with the greatest frequency of *p*-values in the <.05 bin. The observed *p*-curve was significantly flatter than would be expected if the experiments had 33% power, *p* = .0035, indicating an absence of evidential value. Conversely, the half *p*-curve [@Simonsohn2015-mn] was significantly right skewed, suggesting the presence of evidential value. Sensitivity analysis, however, revealed that the half curve does not remain significantly right skewed following removal of the most extreme *p*-value from the sample. \textcolor{red}{The estimated power of the included studies was 5\%, 95\% $CI$ [5\%, 17\%].}

## Interim discussion
The primary results described above suggest that selection effects have caused a seriously distorted record of self-controlled learning. Estimated benefits are less than one third of the naïve estimate, *g* = .107, 95% *CI* \textcolor{red}{[.05, .17]}. The *p*-curve analysis failed to detect robust evidence of a self-controlled learning effect. The performance of the weight-function model depends on the specific conditions present in the meta-analysis, although these conditions are unknowable [@Carter2019-vv]. It was necessary to conduct sensitivity analyses with additional bias correction methods to assess the reliability of the selection-adjusted weight-function model estimate. Based on performance checks conducted under a range of plausible conditions, it was determined that sensitivity analyses conducted with a PEESE meta-regression and *p*-curve effect size estimation would provide good performance coverage across most plausible conditions.

(ref:fig5-caption) *P*-curve analysis of published experiments that were statistically significant at retention. If the included experiments are studying a true null hypothesis the expected distribution of p-values is uniform, represented by the dotted line. If the experiments are studying a true effect, the expected distribution becomes increasingly right skewed as a function of statistical power. The expected right skewed distribution associated with 33% power is plotted by the dashed line. The observed p-curve is plotted by the solid line and was substantially flatter than the 33% power distribution. The half p-curve analysis included p-values below p = .025 and was significantly right skewed. The right skew did not survive deletion of the most extreme value.

```{r fig5, echo = FALSE, fig.cap = "(ref:fig5-caption)", fig.align = "center", out.width="90%"}
knitr::include_graphics("../../figs/fig5.pdf")
```

@@

## Precision-effect with standard error (PEESE) model
When publication bias is present in a body of evidence, sample size and effect size can be negatively correlated [@Stanley2014-xl]. The PEESE model fits a quadratic relationship between effect size and standard error to reflect the intuition that publication bias is stronger for low precision studies than high precision studies. The rationale is that low precision studies need to overestimate effects to achieve significance and get published, while high precision studies can publish without exaggerated effects; thus, creating greater publication bias among lower precision studies [@Carter2019-vv; @Stanley2014-xl]. A weighted-least-squares regression model was fit with effect size regressed on the square of the standard error, weighted by the inverse of the variance:
$$g_{i} = b_{0} + b_{1}se_{i}^2 + e_{i}$$
The PEESE method estimated a non-significant benefit of self-controlled learning after controlling for publication bias, *g* = .054, *p* = .659.

## *P*-curve effect estimation
A *p*-curve model was fit to the overall retention effect size data, unlike the first primary *p*-curve which was fit to the reported significant results. The *p*-curve is a function of sample size and effect size, and because sample size is known, the effect size that provides the best fit to the observed *p*-curve can be estimated [@Simonsohn2014-xq]. A *p*-curve analysis conducted with th`R` package `dmetar` [@Harrer2019-yv] was used to estimate the average effect size among the statistically significant effects in the meta-analysis. The model estimated an average effect of $g$ = .035.[^5] \textcolor{red}{The estimated power of included studies was 7\%, 95\% $CI$ [5\%, 22\%]. Unfortunately, \emph{p}-curve does not perform well in the presence of heterogeneity and these results should be interpreted cautiously.}

[^5]: The *p*-curve of effect sizes was significantly flatter than the expected 33% power curve as well, *p* = .009.

## \textcolor{red}{\emph{Z}-curve}
\textcolor{red}{A \emph{z}-curve was fit to the overall retention data and estimated the power of statistically significant studies (ERR) as 14\%, 95\% CI [5\%, 35\%]. The power of all studies conducted (EDR) was estimated as 6\%, 95\% $CI$ [5\%, 14\%]. The 95\% confidence intervals for both the ERR and EDR failed to include the observed discovery rate of 48\%, suggesting significant publication bias in the data.}
